Appendix S1 - \emph{Estimating carrying capacity for juvenile salmon using quantile random forest models}

### 2020-12-04 analysis/paper/appendix/appendix_s1.Rmd
Implemented revisions suggested by Mike, Sarah and Richie before final submission

---
title: Appendix S1 - _Estimating Carrying Capacity for Juvenile Salmon using Quantile Random Forest Models_
subtitle: "Ecospheres"
author:
- Kevin E. See:
    email: Kevin.See@merck.com
    institute: [biomark]
    correspondence: true
- Michael W. Ackerman:
    email: Mike.Ackerman@merck.com
    institute: [biomark]
    correspondence: false
- Richard A. Carmichael:
    email: Richard.Carmichael@merck.com
    institute: [biomark]
    correspondence: false
- Sarah L. Hoffmann:
    email: Sarah.Hoffmann@merck.com
    institute: [biomark]
    correspondence: false
- Chris Beasley:
    email: Chris.Beasley@merck.com
    institute: [biomark]
    correspondence: false
institute:
  - biomark: Biomark, Inc. 705 South 8th St., Boise, Idaho, 83702, USA
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
    bookdown::pdf_document2:
      fig_caption: yes
      fig_height: 8.5
      fig_width: 6.5
      toc: no
      number_sections: FALSE
      geometry: margin=1in
      latex_engine: "xelatex"
      keep_tex: true
      pandoc_args:
      - --lua-filter=../../templates/scholarly-metadata.lua
      - --lua-filter=../../templates/author-info-blocks2.lua
      - --lua-filter=../../templates/pagebreak.lua
      includes:
        in_header: ../../templates/header_appendixS.tex
    bookdown::word_document2:
      fig_caption: yes
      toc: no
      always_allow_html: true
      number_sections: FALSE
      fig_height: 7
      fig_width: 6
      reference_docx: "../../templates/template.docx" # Insert path for the DOCX file
      pandoc_args:
      - --lua-filter=../../templates/scholarly-metadata.lua
      - --lua-filter=../../templates/author-info-blocks.lua
      - --lua-filter=../../templates/pagebreak.lua
    bookdown::html_document2:
      fig_caption: yes
      fig_height: 8.5
      fig_width: 6.5
      number_sections: FALSE
      pandoc_args:
      - --lua-filter=../../templates/scholarly-metadata.lua
      - --lua-filter=../../templates/author-info-blocks.lua
      - --lua-filter=../../templates/pagebreak.lua
fontsize: 12pt
mainfont: Times New Roman
bibliography:
- ../references.bib
- ../packages.bib
csl: "../../templates/ecological-applications.csl" # Insert path for the bib-style
---

@@

To prevent overfitting, we pared down the more than 100 metrics generated by the CHaMP protocol describing the quantity and quality of fish habitat for each survey site. Habitat metrics were first grouped into broad categories that included channel unit configuration, complexity, fish cover, riparian areas, side channels, stream size, substrate, temperature, water quality, and woody debris. Habitat metrics measuring any large wood volume were scaled by the site length (in 100 m units). To assist in determining the habitat metrics to include in the QRF model, we used the Maximal Information-Based Nonparametric Exploration (MINE) class of statistics [@Reshef2011] to determine those habitat characteristics (covariates) most highly associated with the log of observed parr densities. We calculated the maximal information coefficient (MIC), using the R package *minerva* [@R-minerva], to measure the strength of the linear or non-linear association between the natural log of fish density and each habitat metric [@Reshef2011]. MIC is a measure of correlation that incorporates potential non-linear associations; for example, if there is a quadratic association the MIC value could be high, even when the standard correlation coefficient is low. We excluded categorical variables such as channel type (e.g., meandering, pool-riffle, plane-bed, etc.) because we assumed that other quantitative metrics would capture the differences between those qualitative categorical metrics.

>>>>

---
title: Appendix S1 - _Estimating carrying capacity for juvenile salmon using quantile random forest models_
subtitle: "Ecosphere"
author:
- Kevin E. See:
    email: Kevin.See@merck.com
    institute: [biomark]
    correspondence: true
- Michael W. Ackerman:
    email: Mike.Ackerman@merck.com
    institute: [biomark]
    correspondence: false
- Richard A. Carmichael:
    email: Richard.Carmichael@merck.com
    institute: [biomark]
    correspondence: false
- Sarah L. Hoffmann:
    email: Sarah.Hoffmann@merck.com
    institute: [biomark]
    correspondence: false
- Chris Beasley:
    email: Chris.Beasley@merck.com
    institute: [biomark]
    correspondence: false
institute:
  - biomark: Biomark, Inc. 705 South 8th St., Boise, Idaho, 83702, USA
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
    bookdown::pdf_document2:
      fig_caption: yes
      fig_height: 8.5
      fig_width: 6.5
      toc: no
      number_sections: FALSE
      geometry: margin=1in
      latex_engine: "xelatex"
      keep_tex: true
      pandoc_args:
      - --lua-filter=../../templates/scholarly-metadata.lua
      - --lua-filter=../../templates/author-info-blocks2.lua
      - --lua-filter=../../templates/pagebreak.lua
      includes:
        in_header: ../../templates/header_appendixS.tex
    bookdown::word_document2:
      fig_caption: yes
      toc: no
      always_allow_html: true
      number_sections: FALSE
      fig_height: 7
      fig_width: 6
      reference_docx: "../../templates/template.docx" # Insert path for the DOCX file
      pandoc_args:
      - --lua-filter=../../templates/scholarly-metadata.lua
      - --lua-filter=../../templates/author-info-blocks.lua
      - --lua-filter=../../templates/pagebreak.lua
    bookdown::html_document2:
      fig_caption: yes
      fig_height: 8.5
      fig_width: 6.5
      number_sections: FALSE
      pandoc_args:
      - --lua-filter=../../templates/scholarly-metadata.lua
      - --lua-filter=../../templates/author-info-blocks.lua
      - --lua-filter=../../templates/pagebreak.lua
fontsize: 12pt
mainfont: Times New Roman
bibliography:
- ../references.bib
- ../packages.bib
csl: "../../templates/ecological-applications.csl" # Insert path for the bib-style
---

@@

To prevent overfitting, we pared down the more than 100 metrics generated by the CHaMP protocol [@CHaMP2016] describing the quantity and quality of fish habitat for each survey site. Habitat metrics were first grouped into broad categories that included channel unit configuration, complexity, fish cover, riparian areas, side channels, stream size, substrate, temperature, water quality, and woody debris. Habitat metrics measuring any large wood volume were scaled by the site length (in 100 m units). To assist in determining the habitat metrics to include in the QRF model, we used the Maximal Information-Based Nonparametric Exploration (MINE) class of statistics [@Reshef2011] to determine those habitat characteristics (covariates) most highly associated with the log of observed parr densities. We calculated the maximal information coefficient (MIC), using the R package *minerva* [@R-minerva], to measure the strength of the linear or non-linear association between the natural log of fish density and each habitat metric [@Reshef2011]. MIC is a measure of correlation that incorporates potential non-linear associations; for example, if there is a quadratic association the MIC value could be high, even when the standard correlation coefficient is low. We excluded categorical variables such as channel type (e.g., meandering, pool-riffle, plane-bed, etc.) because we assumed that other quantitative metrics would capture the differences between those qualitative categorical metrics.

### 2020-12-04 analysis/paper/appendix/appendix_s1.tex
updated the author formatting in the appendix so they don't overrun a single line
Implemented revisions suggested by Mike, Sarah and Richie before final submission

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  12pt,
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
  \setmainfont[]{Times New Roman}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Appendix S1 - Estimating Carrying Capacity for Juvenile Salmon using Quantile Random Forest Models},
  pdfauthor={Kevin E. See1,, Michael W. Ackerman1, Richard A. Carmichael1, Sarah L. Hoffmann1, and Chris Beasley1},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs}
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% set figure and table numbering
\usepackage{longtable}
\renewcommand{\thefigure}{S\arabic{figure}}
\setcounter{figure}{0}
\renewcommand{\thetable}{S\arabic{table}}
\setcounter{table}{0}

@@

\title{Appendix S1 - \emph{Estimating Carrying Capacity for Juvenile Salmon using Quantile Random Forest Models}}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{Ecospheres}
\author{Kevin E. See\textsuperscript{1,*}, Michael W. Ackerman\textsuperscript{1}, Richard A. Carmichael\textsuperscript{1}, Sarah L. Hoffmann\textsuperscript{1}, and Chris Beasley\textsuperscript{1}}
\date{December 02, 2020}

@@

To prevent overfitting, we pared down the more than 100 metrics generated by the CHaMP protocol describing the quantity and quality of fish habitat for each survey site. Habitat metrics were first grouped into broad categories that included channel unit configuration, complexity, fish cover, riparian areas, side channels, stream size, substrate, temperature, water quality, and woody debris. Habitat metrics measuring any large wood volume were scaled by the site length (in 100 m units). To assist in determining the habitat metrics to include in the QRF model, we used the Maximal Information-Based Nonparametric Exploration (MINE) class of statistics (Reshef et al. 2011) to determine those habitat characteristics (covariates) most highly associated with the log of observed parr densities. We calculated the maximal information coefficient (MIC), using the R package \emph{minerva} (Filosi et al. 2019), to measure the strength of the linear or non-linear association between the natural log of fish density and each habitat metric (Reshef et al. 2011). MIC is a measure of correlation that incorporates potential non-linear associations; for example, if there is a quadratic association the MIC value could be high, even when the standard correlation coefficient is low. We excluded categorical variables such as channel type (e.g., meandering, pool-riffle, plane-bed, etc.) because we assumed that other quantitative metrics would capture the differences between those qualitative categorical metrics.

@@

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-R-minerva}{}%
Filosi, M., R. Visintainer, and D. Albanese. 2019. Minerva: Maximal information-based nonparametric exploration for variable analysis.

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  12pt,
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
  \setmainfont[]{Times New Roman}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Appendix S1 - Estimating carrying capacity for juvenile salmon using quantile random forest models},
  pdfauthor={Kevin E. See1,, Michael W. Ackerman1, Richard A. Carmichael1, Sarah L. Hoffmann1, and Chris Beasley1},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs}
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% set figure and table numbering
\usepackage{longtable}
\renewcommand{\thefigure}{S\arabic{figure}}
\setcounter{figure}{0}
\renewcommand{\thetable}{S\arabic{table}}
\setcounter{table}{0}

@@

\title{Appendix S1 - \emph{Estimating carrying capacity for juvenile salmon using quantile random forest models}}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{Ecosphere}
\author{Kevin E. See\textsuperscript{1,*}, 
    \and Michael W. Ackerman\textsuperscript{1}, 
    \and Richard A. Carmichael\textsuperscript{1},
    \and Sarah L. Hoffmann\textsuperscript{1}, 
    \and Chris Beasley\textsuperscript{1}}
\date{December 03, 2020}

@@

To prevent overfitting, we pared down the more than 100 metrics generated by the CHaMP protocol (CHaMP 2016) describing the quantity and quality of fish habitat for each survey site. Habitat metrics were first grouped into broad categories that included channel unit configuration, complexity, fish cover, riparian areas, side channels, stream size, substrate, temperature, water quality, and woody debris. Habitat metrics measuring any large wood volume were scaled by the site length (in 100 m units). To assist in determining the habitat metrics to include in the QRF model, we used the Maximal Information-Based Nonparametric Exploration (MINE) class of statistics (Reshef et al. 2011) to determine those habitat characteristics (covariates) most highly associated with the log of observed parr densities. We calculated the maximal information coefficient (MIC), using the R package \emph{minerva} (Filosi et al. 2019), to measure the strength of the linear or non-linear association between the natural log of fish density and each habitat metric (Reshef et al. 2011). MIC is a measure of correlation that incorporates potential non-linear associations; for example, if there is a quadratic association the MIC value could be high, even when the standard correlation coefficient is low. We excluded categorical variables such as channel type (e.g., meandering, pool-riffle, plane-bed, etc.) because we assumed that other quantitative metrics would capture the differences between those qualitative categorical metrics.

@@

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-CHaMP2016}{}%
CHaMP. 2016. Scientific protocol for salmonid habitat surveys within the Columbia habitat monitoring program. Bonneville Power Administration Portland, Oregon, USA.

\leavevmode\hypertarget{ref-R-minerva}{}%
Filosi, M., R. Visintainer, and D. Albanese. 2019. Minerva: Maximal information-based nonparametric exploration for variable analysis.
